---
title: "Geneset Growth Stats"
output: html_document
params:
  species: Human
  identifier: entrezgene
  year: 2019
  width: 600
  height: 600
  pointsz: 14
---

# Changes/Growth of Genesets By Source over the years
```{r load libraries, echo=FALSE,message=FALSE, warning=FALSE}

## Load in required libraries

tryCatch(expr = { library("rmarkdown")}, 
         error = function(e) {  install.packages("rmarkdown")}, 
         finally = library("rmarkdown"))

#install required R and bioconductor packages
tryCatch(expr = { library("RCurl")}, 
         error = function(e) {  install.packages("RCurl")}, 
         finally = library("RCurl"))

tryCatch(expr = { library("ggplot2")}, 
         error = function(e) {  install.packages("ggplot2")}, 
         finally = library("ggplot2"))

```


```{r include=FALSE}
#Download all the directory names from the downloads website.

download_url = "http://download.baderlab.org/EM_Genesets/"

#list all the files on the server
filenames = getURL(download_url)
tc = textConnection(filenames)
contents = readLines(tc)
close(tc)

rx = gregexpr('(?<=<a href=\")(.*_.*_.*)(?=\">)',contents, perl = TRUE)
download_directories = unlist(regmatches(contents, rx))
download_directories <- download_directories [grep(download_directories,pattern = ".php",invert=TRUE)]
```


```{r message=FALSE, warning=FALSE,include=FALSE}

get_summary_file <- function(species = "Human", identifier="entrezgene", year=2014) {
      
      
      stats_summary_table <- c()
      stats_summary_ggplot <- c()
      
      #get the summary files for each directory 
      for(i in 1:length(download_directories )){
             #only go back as far as December 2014
          if(length(grep(download_directories[i], pattern=year )>0)){
            break
          }
              current_directory = paste(download_url,download_directories [i], species,"/",sep="")
            
            #list all the files in this directory
            current_filenames = getURL(current_directory)
            tc = textConnection(current_filenames)
            contents = readLines(tc)
            close(tc)
            
            rx = gregexpr("(?<=<a href=\")(Summary_Gene.*.)(.txt)(?=\">)",
              contents, perl = TRUE)
            log_files = unlist(regmatches(contents, rx))
            
            current_log_file <- log_files[grep(log_files,pattern=identifier)]
            
            if(length(current_log_file) == 0 && length(log_files == 1)){
              current_log_file = log_files
            }
            
            dest_log_file <- file.path(getwd(),current_log_file )
            
            download.file(
                paste(current_directory,current_log_file,sep=""),
                destfile=dest_log_file,quiet = TRUE
            )
            
            current_log_contents <- read.table(dest_log_file,header=FALSE,sep="\t",stringsAsFactors = FALSE,strip.white = TRUE)
            current_log_contents <- current_log_contents[grep(current_log_contents[,1],pattern=".gmt"),]
            #current_log_contents <- sapply(current_log_contents,FUN = function(x){strsplit(x,pattern=" ")})
            
            stats <- data.frame(gmt_file = tolower(sapply(current_log_contents,FUN = function(x){unlist(strsplit(x,split=" ")[[1]][2])})),
                           num_sets = as.integer(sapply(current_log_contents,FUN = function(x){unlist(strsplit(x,split=" ")[[1]][1])})))
            
            rownames(stats) <- c()
            
            #replace the dates with nothing so they are comparable
            stats[,1] <- gsub(stats[,1],pattern = tolower(substr(download_directories [i],1,nchar(download_directories [i])-1)),replacement = "")
            
            #make sure that stats table is unique (there is an issue with multiple rows being added)
            stats <- aggregate(stats, by=list(stats$gmt_file),FUN=max)
            
            #get the date of this data
            proper_date <- as.Date(download_directories [i],"%B_%d_%Y")
            
            #colnames(stats)[2] <- proper_date
      
            
            if(i == 1){
      
              stats_summary_ggplot <- data.frame(stats, date = proper_date)
              colnames(stats)[2] <- download_directories [i]
              stats_summary_table <- stats
            } else{
              stats_summary_ggplot <- rbind(stats_summary_ggplot, 
                                            data.frame(stats, date = proper_date))
              colnames(stats)[2] <- download_directories [i]
              merged_data <- merge(stats_summary_table, stats,by.x = 1, by.y=1,
                                   all = TRUE)
              stats_summary_table <- merged_data
              
            }
      }
      
      return(stats_summary_ggplot)
}
```


```{r message=FALSE, warning=FALSE,include=FALSE}
get_identifier_conversion_file <- function(species = "Human", identifier="entrezgene", year=2014) {
  
  #filter the download directories by the specified year
  
  
  ids_summary_table <- c()
  ids_summary_ggplot <- c()
  
  #get the summary files for each directory 
  for(i in 1:length(download_directories )){
    #only go back as far as December 2014
    #if(length(grep(download_directories[i], pattern=year )>0)){
    if(unlist(strsplit(gsub(download_directories[i],pattern = "/",replacement = ""),split = "_"))[3] < year) {
      next
    }
    current_directory = paste(download_url,download_directories[i], species,"/",sep="")
    
    #list all the files in this directory
    current_filenames = getURL(current_directory)
    tc = textConnection(current_filenames)
    contents = readLines(tc)
    close(tc)
    
    rx = gregexpr("(?<=<a href=\")(Build_summary.*.)(.log)(?=\">)",
                  contents, perl = TRUE)
    log_files = unlist(regmatches(contents, rx))
    
    current_log_file <- log_files[grep(log_files,pattern=identifier)]
    
    if(length(current_log_file) == 0 && length(log_files == 1)){
      current_log_file = log_files
    }
    
    dest_log_file <- file.path(getwd(),current_log_file )
    
    if(length(dest_log_file) == 0){
      break
    }
    
    download.file(
      paste(current_directory,current_log_file,sep=""),
      destfile=dest_log_file,quiet = TRUE
    )
    
    current_log_contents <- read.table(dest_log_file,header=FALSE,sep="\t",stringsAsFactors = FALSE,strip.white = TRUE,fill = TRUE)
    colnames(current_log_contents) <- c("empty","Filename",	"Original Identifier",	"ID translated",	"Total #genes",	 "#genes translated",	"#genes missing",	 "percent not translated",	 "Total Annotations",	 "Annotations not translated",	 "percent annot not translated","empty")
    
    #the first entry in every file is WikiPathways and it is missing the empty column at the front
    current_log_contents[2,2:11] <- current_log_contents[2,1:10]
    
    current_log_contents <- current_log_contents[grep(current_log_contents[,2],pattern=".gmt"),]
    current_log_contents[,2] <- tolower(current_log_contents[,2])
    stats<- current_log_contents[,c(2:11)]
    
    
    #get the date of this data
    proper_date <- as.Date(download_directories [i],"%B_%d_%Y")
    
    #colnames(stats)[2] <- proper_date
    
    
    if(i == 1){
      
      ids_summary_ggplot <- data.frame(stats, date = proper_date)
      #colnames(stats)[2] <- download_directories [i]
      #ids_summary_table <- stats
    } else{
      ids_summary_ggplot <- rbind(ids_summary_ggplot, 
                                  data.frame(stats, date = proper_date))
      #colnames(stats)[2] <- download_directories [i]
      #merged_data <- merge(ids_summary_table, stats,by.x = 1, by.y=1,
      #                     all = TRUE)
      #ids_summary_table <- merged_data
      
    }
  }
  
  return(ids_summary_ggplot)
}
```



```{r message=FALSE, warning=FALSE,include=FALSE}

get_identifier_conversion_file_by_type <- function(species = "Human", identifier="entrezgene", year=2014, type="Pathways") {
  
  #filter the download directories by the specified year
  ids_summary_table <- c()
  ids_summary_ggplot <- c()
  
  #get the summary files for each directory 
  for(i in 1:length(download_directories )){
    #only go back as far as December 2014
    #if(length(grep(download_directories[i], pattern=year )>0)){
    if(unlist(strsplit(gsub(download_directories[i],pattern = "/",replacement = ""),split = "_"))[3] < year) {
      next
    }
    current_directory = paste(download_url,download_directories[i], species,"/",identifier,"/",type,"/",sep="")
    
    #list all the files in this directory
    current_filenames = getURL(current_directory)
    tc = textConnection(current_filenames)
    contents = readLines(tc)
    close(tc)
    
    rx = gregexpr("(?<=<a href=\")(.*.translation.*.)(.log)(?=\">)",
                  contents, perl = TRUE)
    log_files = unlist(regmatches(contents, rx))
    
    if(is.na(log_files) || length(log_files) ==0){
      next
    }
    
    #get the date of this data
    proper_date <- as.Date(download_directories[i],"%B_%d_%Y")
    
    #for each of the log files
    for(j in 1:length(log_files)){
        db_name <- unlist(strsplit(log_files[j],split = "_"))[2]
      dest_log_file <- file.path(getwd(),log_files[j] )
      download.file(
        paste(current_directory,log_files[j],sep=""),
        destfile=dest_log_file,quiet = TRUE
      )
      current_log_contents <- read.table(dest_log_file,header=FALSE,sep="\t",
                                         stringsAsFactors = FALSE,strip.white = TRUE,fill = TRUE)
     
      num <- 1
      #if there is more than one stats record for this file
      if((dim(current_log_contents)[1] >  12) && (proper_date > as.Date("2020-11-17"))){
        stats <- t(current_log_contents[1:10,2])
        while(num*11 < dim(current_log_contents)[1]){
          stats <- rbind(stats, current_log_contents[(11*num+1):(11*num+10),2])
          num <- num+1
        }
      } else if((dim(current_log_contents)[1] >  12) && (proper_date <= as.Date("2020-11-17"))){
        stats <- cbind(t(current_log_contents[1:4,2]), "",t(current_log_contents[5:9,2]))
        while(num*10 < dim(current_log_contents)[1]){
          start <- (10*num+1)
          end <- (10*num+9)
          stats <- rbind(stats, cbind(t(current_log_contents[start:(start + 4),2]),"",
                         t(current_log_contents[(start+5):end,2])))
          num <- num+1
        }
      } else if(proper_date <= as.Date("2020-11-17")){
        stats <- cbind(t(current_log_contents[1:4,2]), "",t(current_log_contents[5:9,2]))
        } else{
         stats <- t(current_log_contents[1:10,2])
      }
      
      
      
      if(i == 1 && j == 1){
        ids_summary_ggplot <- data.frame(source=db_name,stats, date = proper_date)
      } else{
        ids_summary_ggplot <- rbind(ids_summary_ggplot, 
                                    data.frame(source=db_name, stats, date = proper_date))
        
      }
      
      
    }
    
    
  }
  
  colnames(ids_summary_ggplot) <- c("source","Filename",	"Original Identifier",	"ID translated",	"Total #genes",	 "#genes translated",	"#genes missing",	 "percent not translated",	 "Total Annotations",	 "Annotations not translated",	 "percent annot not translated","date")
  
  return(ids_summary_ggplot)
}

```







```{r  message=FALSE, warning=FALSE,include=FALSE}
library(ggplot2)

species = params$species
identifier = params$identifier
year = params$year

#make a dir for the current figures
figure_directory <- file.path( "figures",paste(species, identifier,year,
"to",format(Sys.Date(),"%B_%Y"),sep="_"))

if(!dir.exists(file.path(getwd(),figure_directory))){
	dir.create(file.path(getwd(),figure_directory))
}

stats_summary_ggplot <- get_summary_file(species,identifier,year)

overall_figure_filename <- "Overall_geneset_stats.png"
png(file.path(getwd(), figure_directory, overall_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

ggplot(stats_summary_ggplot, aes(x=date, y=num_sets, group=gmt_file,color=gmt_file)) +
    geom_line() +  
        theme(legend.position = "right", 
        legend.key.size = unit(0.25, 'cm'), #change legend key size
        legend.key.height = unit(0.25, 'cm'), #change legend key height
        legend.key.width = unit(0.25, 'cm'), #change legend key width
        legend.title = element_text(size=9), #change legend title font size
        legend.direction = "vertical",
         legend.spacing = unit(0.04, "cm"),
        legend.text = element_text(size=10)) + #change legend text font size
       ggtitle(paste("Number of genesets over time for", species,"from", year, "- Present", sep = " " )) +
  guides(colour = guide_legend(ncol = 1)) + 
    ylab("Number of genesets") + xlab("time")

dev.off()
```

![](`r file.path(".",figure_directory, overall_figure_filename)`)


```{r  message=FALSE, warning=FALSE,include=FALSE}
subset_go <- stats_summary_ggplot[grep(stats_summary_ggplot[,1],pattern = "go"),]

GO_figure_filename <- "GO_geneset_stats.png"

png(file.path(getwd(),figure_directory,GO_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

ggplot(subset_go, aes(x=date, y=num_sets, group=gmt_file,color=gmt_file)) +
    geom_line() + theme(legend.position = "right", 
        legend.key.size = unit(0.25, 'cm'), #change legend key size
        legend.key.height = unit(0.25, 'cm'), #change legend key height
        legend.key.width = unit(0.25, 'cm'), #change legend key width
        legend.title = element_text(size=9), #change legend title font size
        legend.direction = "vertical",
         legend.spacing = unit(0.04, "cm"),
        legend.text = element_text(size=10)) + #change legend text font size  
       ggtitle(paste("Number of GO genesets for", species,"from", year, "- Present", sep = " " )) +
  guides(colour = guide_legend(ncol = 1)) + 
    ylab("Number of genesets")  + xlab("time")
dev.off()
```

![](`r file.path(".",figure_directory, GO_figure_filename)`)

```{r  message=FALSE, warning=FALSE,include=FALSE}
subset_drugbank <- stats_summary_ggplot[grep(stats_summary_ggplot[,1],pattern = "drug|phenotype|mirs|transcription"),] 

Other_figure_filename <- "Other_geneset_stats.png"

png(file.path(getwd(),figure_directory, Other_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)


ggplot(subset_drugbank, aes(x=date, y=num_sets, group=gmt_file,color=gmt_file)) +
    geom_line() + theme(legend.position = "right", 
        legend.key.size = unit(0.25, 'cm'), #change legend key size
        legend.key.height = unit(0.25, 'cm'), #change legend key height
        legend.key.width = unit(0.25, 'cm'), #change legend key width
        legend.title = element_text(size=9), #change legend title font size
        legend.direction = "vertical",
         legend.spacing = unit(0.04, "cm"),
        legend.text = element_text(size=10)) + #change legend text font size   
        ggtitle(paste("Number of Phenotype genesets over time for", species,"from", year, "- Present", sep = " " )) +
  guides(colour = guide_legend(ncol = 1)) + 
    ylab("Number of genesets") + xlab("time")
dev.off()
```


![](`r file.path(".",figure_directory, Other_figure_filename)`)

```{r  message=FALSE, warning=FALSE,include=FALSE}
subset_pathways <- stats_summary_ggplot[grep(stats_summary_ggplot[,1],pattern = "drug|phenotype|mirs|go|transcription", invert = TRUE),] 

Pathway_figure_filename <- "Pathway_geneset_stats.png"

png(file.path(getwd(),figure_directory, Pathway_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

ggplot(subset_pathways, aes(x=date, y=num_sets, group=gmt_file,color=gmt_file)) +
    geom_line() + theme(legend.position = "right", 
        legend.key.size = unit(0.25, 'cm'), #change legend key size
        legend.key.height = unit(0.25, 'cm'), #change legend key height
        legend.key.width = unit(0.25, 'cm'), #change legend key width
        legend.title = element_text(size=9), #change legend title font size
        legend.direction = "vertical",
         legend.spacing = unit(0.04, "cm"),
        legend.text = element_text(size=10)) + #change legend text font size  
    ggtitle(paste("Number of Pathway genesets over time for", species,"from", year, "- Present", sep = " " )) +
  guides(colour = guide_legend(ncol = 1)) + 
    ylab("Number of genesets") + xlab("time")
dev.off()
```

![](`r file.path(".",figure_directory, Pathway_figure_filename)`)

# Changes/Growth of Identifier mapping By Source over the years

```{r message=FALSE, warning=FALSE,include=FALSE}

species = params$species
identifier = "symbol"
year = "2019"

#make a dir for the current figures
figure_directory <- file.path( "figures",paste(species, identifier,year,
"to",format(Sys.Date(),"%B_%Y"),sep="_"))

if(!dir.exists(file.path(getwd(),figure_directory))){
	dir.create(file.path(getwd(),figure_directory))
}

```


```{r message=FALSE, warning=FALSE,include=FALSE}
id_summary_ggplot_pathways <- get_identifier_conversion_file_by_type(species,identifier,"2019","Pathways")

subset_id_summary <- data.frame(source = id_summary_ggplot_pathways$source, not_translated = as.numeric(id_summary_ggplot_pathways$`Annotations not translated`), total = as.numeric(id_summary_ggplot_pathways$`Total Annotations`), date= id_summary_ggplot_pathways$date, stringsAsFactors = FALSE)

subset_id_summary_agg <- aggregate(subset_id_summary[,2:3],by = list(subset_id_summary$source,subset_id_summary$date),FUN = sum)
colnames(subset_id_summary_agg) <- c("source","date","not_translated","total")

Pathwayid_figure_filename <- "Pathway_identifier_stats.png"

png(file.path(getwd(),figure_directory, Pathwayid_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

  ggplot(subset_id_summary_agg, aes(x=date, y=(not_translated/total)*100, group=source,color=source)) +
    geom_line() + theme(legend.position = "right",
    legend.key.size = unit(0.25, 'cm'), #change legend key size
    legend.key.height = unit(0.25, 'cm'), #change legend key height
    legend.key.width = unit(0.25, 'cm'), #change legend key width
    legend.title = element_text(size=9), #change legend title font size
    legend.direction = "vertical",
    legend.spacing = unit(0.04, "cm"),
    legend.text = element_text(size=10)) + #change legend text font size
    ggtitle(paste("Percent missing Annotation translations  \n for Pathways", species,"from", year, "- Present", sep = " " )) +
    guides(colour = guide_legend(ncol = 1)) +
    ylab("Percent (1-100%) Annotations not translated") + xlab("time") 

dev.off()

```

![](`r file.path(".",figure_directory, Pathwayid_figure_filename)`)

There were issues with Panther data from 02-2019 until 11-2020.  To get a better view of the identifier mapping for the remaining pathway databases exclude panther data. 

```{r message=FALSE, warning=FALSE,include=FALSE}
id_summary_ggplot_pathways <- get_identifier_conversion_file_by_type(species,identifier,"2019","Pathways")

subset_id_summary <- data.frame(source = id_summary_ggplot_pathways$source, not_translated = as.numeric(id_summary_ggplot_pathways$`Annotations not translated`), total = as.numeric(id_summary_ggplot_pathways$`Total Annotations`), date= id_summary_ggplot_pathways$date, stringsAsFactors = FALSE)

subset_id_summary_agg <- aggregate(subset_id_summary[,2:3],by = list(subset_id_summary$source,subset_id_summary$date),FUN = sum)
colnames(subset_id_summary_agg) <- c("source","date","not_translated","total")

subset_id_summary_agg <- subset_id_summary_agg[which(subset_id_summary_agg$not_translated/subset_id_summary_agg$total < 0.25),]

Pathwayid2_figure_filename <- "Pathway_nopanther_identifier_stats.png"

png(file.path(getwd(),figure_directory, Pathwayid2_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

  ggplot(subset_id_summary_agg, aes(x=date, y=(not_translated/total)*100, group=source,color=source)) +
    geom_line() + theme(legend.position = "right",
    legend.key.size = unit(0.25, 'cm'), #change legend key size
    legend.key.height = unit(0.25, 'cm'), #change legend key height
    legend.key.width = unit(0.25, 'cm'), #change legend key width
    legend.title = element_text(size=9), #change legend title font size
    legend.direction = "vertical",
    legend.spacing = unit(0.04, "cm"),
    legend.text = element_text(size=10)) + #change legend text font size
    ggtitle(paste("Percent missing Annotation translations  \n for Pathways", species,"from", year, "- Present", sep = " " )) +
    guides(colour = guide_legend(ncol = 1)) +
    ylab("Percent (1-100%) Annotations not translated") + xlab("time") 

dev.off()

```

![](`r file.path(".",figure_directory, Pathwayid2_figure_filename)`)



```{r message=FALSE, warning=FALSE,include=FALSE}
id_summary_ggplot_DrugTargets <- get_identifier_conversion_file_by_type(species,identifier,"2021","DrugTargets")

subset_id_summary <- data.frame(source = id_summary_ggplot_DrugTargets$Filename, not_translated = as.numeric(id_summary_ggplot_DrugTargets$`Annotations not translated`), total = as.numeric(id_summary_ggplot_DrugTargets$`Total Annotations`), date= id_summary_ggplot_DrugTargets$date, stringsAsFactors = FALSE)

subset_id_summary_agg <- aggregate(subset_id_summary[,2:3],by = list(subset_id_summary$source,subset_id_summary$date),FUN = sum)
colnames(subset_id_summary_agg) <- c("Filename","date","not_translated","total")

Drugid_figure_filename <- "Drugbank_identifier_stats.png"

png(file.path(getwd(),figure_directory, Drugid_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

ggplot(subset_id_summary_agg, aes(x=date, y=(not_translated/total)*100, group=Filename,color=Filename)) +
  geom_line() + theme(legend.position = "right",
  legend.key.size = unit(0.25, 'cm'), #change legend key size
  legend.key.height = unit(0.25, 'cm'), #change legend key height
  legend.key.width = unit(0.25, 'cm'), #change legend key width
  legend.title = element_text(size=9), #change legend title font size
  legend.direction = "vertical",
  legend.spacing = unit(0.04, "cm"),
  legend.text = element_text(size=10)) + #change legend text font size
  ggtitle(paste("Percent missing Annotation translations  \n for DrugTargets", species,"from", year, "- Present", sep = " " )) +
  guides(colour = guide_legend(ncol = 1)) +
  ylab("Percent (1-100%) Annotations not translated") + xlab("time") 

dev.off()

```
![](`r file.path(".",figure_directory, Drugid_figure_filename)`)



```{r message=FALSE, warning=FALSE,include=FALSE}
id_summary_ggplot_miRs <- get_identifier_conversion_file_by_type(species,identifier,"2019","miRs")

subset_id_summary <- data.frame(source = id_summary_ggplot_miRs$source, not_translated = as.numeric(id_summary_ggplot_miRs$`Annotations not translated`), total = as.numeric(id_summary_ggplot_miRs$`Total Annotations`), date= id_summary_ggplot_miRs$date, stringsAsFactors = FALSE)

subset_id_summary_agg <- aggregate(subset_id_summary[,2:3],by = list(subset_id_summary$source,subset_id_summary$date),FUN = sum)
colnames(subset_id_summary_agg) <- c("source","date","not_translated","total")

Mirid_figure_filename <- "Mir_identifier_stats.png"

png(file.path(getwd(),figure_directory, Mirid_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

  ggplot(subset_id_summary_agg, aes(x=date, y=(not_translated/total)*100, group=source,color=source)) +
    geom_line() + theme(legend.position = "right",
    legend.key.size = unit(0.25, 'cm'), #change legend key size
    legend.key.height = unit(0.25, 'cm'), #change legend key height
    legend.key.width = unit(0.25, 'cm'), #change legend key width
    legend.title = element_text(size=9), #change legend title font size
    legend.direction = "vertical",
    legend.spacing = unit(0.04, "cm"),
    legend.text = element_text(size=10)) + #change legend text font size
    ggtitle(paste("Percent missing Annotation translations  \n for miRs", species,"from", year, "- Present", sep = " " )) +
    guides(colour = guide_legend(ncol = 1)) +
    ylab("Percent Annotations not translated") + xlab("time") 

dev.off()

```
![](`r file.path(".",figure_directory, Mirid_figure_filename)`)

```{r message=FALSE, warning=FALSE,include=FALSE}
id_summary_ggplot_DiseasePhenotypes <- get_identifier_conversion_file_by_type(species,identifier,"2021","DiseasePhenotypes")

subset_id_summary <- data.frame(source = id_summary_ggplot_DiseasePhenotypes$source, not_translated = as.numeric(id_summary_ggplot_DiseasePhenotypes$`Annotations not translated`), total = as.numeric(id_summary_ggplot_DiseasePhenotypes$`Total Annotations`), date= id_summary_ggplot_DiseasePhenotypes$date, stringsAsFactors = FALSE)

subset_id_summary_agg <- aggregate(subset_id_summary[,2:3],by = list(subset_id_summary$source,subset_id_summary$date),FUN = sum)
colnames(subset_id_summary_agg) <- c("source","date","not_translated","total")

Diseaseid_figure_filename <- "Disease_identifier_stats.png"

png(file.path(getwd(),figure_directory, Diseaseid_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

  ggplot(subset_id_summary_agg, aes(x=date, y=(not_translated/total)*100, group=source,color=source)) +
    geom_line() + theme(legend.position = "right",
    legend.key.size = unit(0.25, 'cm'), #change legend key size
    legend.key.height = unit(0.25, 'cm'), #change legend key height
    legend.key.width = unit(0.25, 'cm'), #change legend key width
    legend.title = element_text(size=9), #change legend title font size
    legend.direction = "vertical",
    legend.spacing = unit(0.04, "cm"),
    legend.text = element_text(size=10)) + #change legend text font size
    ggtitle(paste("Percent missing Annotation translations \n for DiseasePhenotypes", species,"from", year, "- Present", sep = " " )) +
    guides(colour = guide_legend(ncol = 1)) +
    ylab("Percent Annotations not translated") + xlab("time") 

dev.off()
```

![](`r file.path(".",figure_directory, Diseaseid_figure_filename)`)

```{r message=FALSE, warning=FALSE,include=FALSE}
#go through each unique file and graph the percentage of ids that are not mapped
id_summary_ggplot <- get_identifier_conversion_file(species,identifier,year)

subset_pathways <- id_summary_ggplot[grep(id_summary_ggplot[,1],pattern = "drug|phenotype|mirs|go|transcription", invert = TRUE),] 

#collapse all the pathbank files
subset_id_summary <- id_summary_ggplot[,c("Filename","Annotations.not.translated","Total.Annotations","date")]

#divide out any troublesome files that have more than 50% identifier loss
files_with_issues <- subset_id_summary[which(as.numeric(subset_id_summary$Annotations.not.translated)/as.numeric(subset_id_summary$Total.Annotations) > 0.5),]

#remove the files with issues
subset_id_summary <- subset_id_summary[which(as.numeric(subset_id_summary$Annotations.not.translated)/as.numeric(subset_id_summary$Total.Annotations) <= 0.5),]

#the pathbank files are all together - collapse them to just one file
pw_subset <- subset_id_summary[grep(subset_id_summary$Filename,pattern="^pw"),]
pw_subset[,2] <- as.numeric(pw_subset[,2])
pw_subset[,3] <- as.numeric(pw_subset[,3])
subset_id_summary <- subset_id_summary[grep(subset_id_summary$Filename,pattern="^pw",invert = TRUE),]
pw_aggregated <- cbind("Pathbank",aggregate(pw_subset[,2:3],by = list(pw_subset$date),FUN = sum)[,c(2,3,1)])
colnames(pw_aggregated) <- colnames(subset_id_summary)
subset_id_summary <- rbind(subset_id_summary,pw_aggregated)

#misc pathways - get all the owl files. 
misc_subset <- subset_id_summary[grep(subset_id_summary$Filename,pattern=".owl"),]
#remove them from the subset_id_summary
subset_id_summary <- subset_id_summary[grep(subset_id_summary$Filename,pattern="owl",invert = TRUE),]
#add back in reactome, nci-nature, and human cyc
subset_id_summary <- rbind(subset_id_summary,misc_subset[grep(misc_subset$Filename,pattern="homo_sapiens"),])
misc_subset <- misc_subset[grep(misc_subset$Filename,pattern="homo_sapiens",invert = TRUE),]
subset_id_summary <- rbind(subset_id_summary,misc_subset[grep(misc_subset$Filename,pattern="nci-nature"),])
misc_subset <- misc_subset[grep(misc_subset$Filename,pattern="nci-nature",invert = TRUE),]
subset_id_summary <- rbind(subset_id_summary,misc_subset[grep(misc_subset$Filename,pattern="biopax-level"),])
misc_subset <- misc_subset[grep(misc_subset$Filename,pattern="biopax-level",invert = TRUE),]

#collapse the remaining owl files by date into Misc biopax pathways.
misc_subset[,2] <- as.numeric(misc_subset[,2])
misc_subset[,3] <- as.numeric(misc_subset[,3])
misc_aggregated <- cbind("Panther",aggregate(misc_subset[,2:3],by = list(misc_subset$date),FUN = sum)[,c(2,3,1)])
colnames(misc_aggregated) <- colnames(subset_id_summary)
subset_id_summary <- rbind(subset_id_summary,misc_aggregated)


#shorten the names
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="biopax-level")] <- "HumanCyc"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="wikipathways")] <- "WikiPathways"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="homo_sapiens")] <- "Reactome"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="nci-nature")] <- "NCI-nature"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="go_mf_with")] <- "GO_MF_withIEA"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="go_mf_no")] <- "GO_MF_noIEA"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="go_cc_with")] <- "GO_CC_withIEA"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="go_cc_no")] <- "GO_CC_noIEA"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="go_bp_with")] <- "GO_BP_withIEA"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="go_bp_no")] <- "GO_BP_noIEA"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="drugbank_all")] <- "drugbank_all"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="drugbank_approved")] <- "drugbank_approved"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="drugbank_exp")] <- "drugbank_exp"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="drugbank_nut")] <- "drugbank_nut"
subset_id_summary$Filename[grep(subset_id_summary$Filename,pattern="kegg")] <- "Kegg"

subset_id_summary[,2] <- as.numeric(subset_id_summary[,2])
subset_id_summary[,3] <- as.numeric(subset_id_summary[,3])

#limit to just GO
subset_id_summary <- subset_id_summary[grep(subset_id_summary$Filename,pattern="GO_"),]


GOid_figure_filename <- "GO_identifier_stats.png"

png(file.path(getwd(),figure_directory, GOid_figure_filename),
    width = params$width, height = params$height,  pointsize = params$pointsz)

ggplot(subset_id_summary, aes(x=date, y=(Annotations.not.translated/Total.Annotations)*100, group=Filename,color=Filename)) +
    geom_line() + theme(legend.position = "right", 
        legend.key.size = unit(0.25, 'cm'), #change legend key size
        legend.key.height = unit(0.25, 'cm'), #change legend key height
        legend.key.width = unit(0.25, 'cm'), #change legend key width
        legend.title = element_text(size=9), #change legend title font size
        legend.direction = "vertical",
         legend.spacing = unit(0.04, "cm"),
        legend.text = element_text(size=10)) + #change legend text font size  
    ggtitle(paste("Percent missing Annotation translations for", species,"from", year, "- Present", sep = " " )) + 
  guides(colour = guide_legend(ncol = 1)) + 
    ylab("Percent Annotations not translated") + xlab("time")

dev.off()


```



![](`r file.path(".",figure_directory, GOid_figure_filename)`)


